---
title: "Multilingual LLM Evaluation: Idiom Interpretation and Translation"
author: "Nabiha Chowdhury"
date: "`r Sys.Date()`"
output:
  html_document: default
---


## Table #1
```{r}
library(tibble)

researcher <- rep("Nabiha", 36)

llm_type <- c(rep("DuckAI:GPT5", 9), rep("Claude", 9),rep("ChatGPT", 9), rep("Gemini", 9))

idiom <- c(rep("Its raining cats and dogs", 3), rep("Don’t cry over spilled milk", 3), rep("The Whole Nine Yards", 3), rep("Its raining cats and dogs", 3), rep("Don’t cry over spilled milk", 3), rep("The Whole Nine Yards", 3), rep("Its raining cats and dogs", 3), rep("Don’t cry over spilled milk", 3), rep("The Whole Nine Yards", 3), rep("Its raining cats and dogs", 3), rep("Don’t cry over spilled milk", 3), rep("The Whole Nine Yards", 3))

language <- c("English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese","English", "Arabic", "Chinese")

llm_judge <- c(rep("MetaAI and Copilot", 36))

origin <- c(5,3,4.5,5,5,5,5,5,5,4,5,5,5,4,4,3,5,4,4.5,5,5,5,5,4.5,5,5,5,3.5,4.5,3.5,5,5,5,5,3.5,3.5)

accuracy <- c(5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5)

score <- (origin + accuracy) / 2

llms_data <- tibble(
  researcher = researcher,         # Nabiha
  llm = llm_type,                  # ChatGPT, DuckAI, Gemini, Claude
  idiom_prompt = idiom,            # idiom used
  language_type = language,        # language used
  llm_judge = llm_judge,           # llm used to judge
  origin_accuracy = origin,        # correct origin
  translation_accuracy = accuracy, # correct translation
  overall_score = score)           # score of llm's
```


## Table #2
```{r}

name <- c("Copilot", "MetaAI", "Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI", "Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI","Copilot", "MetaAI")

idiom_type <- c(rep("Its raining cats and dogs", 6), rep("Don’t cry over spilled milk", 6), rep("The Whole Nine Yards", 6), rep("Its raining cats and dogs", 6), rep("Don’t cry over spilled milk", 6), rep("The Whole Nine Yards", 6), rep("Its raining cats and dogs", 6), rep("Don’t cry over spilled milk", 6), rep("The Whole Nine Yards", 6), rep("Its raining cats and dogs", 6), rep("Don’t cry over spilled milk", 6), rep("The Whole Nine Yards", 6))

llm_used <- c(rep("DuckAI", 18), rep("Claude", 18), rep("ChatGPT", 18), rep("Gemini", 18))

language_used <- c("English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese", "English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese","English", "English", "Arabic","Arabic", "Chinese", "Chinese")

origin_results <- c(5,5,1,5,4,5,5,5,5,5,5,5,5,5,5,5,5,5,4,4,5,5,5,5,5,5,5,3,4,4,3,3,5,5,4,4, 
                    4,5,5,5,5,5,5,5,5,5,4,5,5,5,5,5,5,5,
                    4,3,4,5,4,3,5,5,5,5,5,5,5,5,2,5,4,3)

translation_results <- c(5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5, 
                         5,5,5,5,5,5,5,5,5,5,4,5,5,5,5,5,5,5,
                         5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5)

average_score <- ((origin_results + translation_results)/2)


llm_scores <- tibble(
  LLM_Judge = name,
  Idiom_prompt = idiom_type,
  LLM_used_for_original_prompt = llm_used,
  Language = language_used,
  Origin_score = origin_results,
  Translation_score = translation_results,
  Overall_score = average_score)

head(llm_scores, 5)
```


## ANOVA testing 
```{r}
# anova score vs languages
summary(aov(overall_score ~ language_type, data = llms_data))
```

HO: there is no difference in the overall score depending on which language of the three is used for the idiom translation

The resulting pvalue 0.94 > 0.05, which is not statically significant so we fail to reject the null hypothesis, there is NO difference in the overall score depending on which language of the three is used for the idiom translation

```{r}
# anova score vs llm used 
summary(aov(Overall_score ~ LLM_used_for_original_prompt, data = llm_scores))
```
HO: there is no difference in the overall score depending on which LLM is used to ask for translation/origin of idioms. 

The resulting pvalue 0.114 > 0.05, which is not statically significant so we fail to reject the null hypothesis, there is NO difference in the overall score when using different LLM's to translate idioms

## Visualizations 
```{r}
library(tidyverse)
#  Average Overall Score by LLM
ggplot(llms_data, aes(x = llm, y = overall_score, fill = language_type)) +
  geom_boxplot() +
  labs(
    title="LLM vs Overall Score")
```

```{r}
#  Average Overall Score by Language
ggplot(llms_data, aes(x = language_type, y = overall_score, fill = language_type)) +
  stat_summary(geom = "bar") +
  facet_wrap(~ idiom_prompt) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "Average Overall Score by Language", y = "Average Score", x = "Language")
```

